---
title: "Jun Pan 607 Project 2 dataset 1"
author: "Jun Pan"
date: "October 5, 2018"
output: html_document
---

New York City Leading Causes of Death 
from https://data.cityofnewyork.us/Health/New-York-City-Leading-Causes-of-Death/jb7j-dtam
down load csv file to local computer for data analysis

set environment
```{r}
library(dplyr)
library(knitr)
library(DT)
options(dplyr.width = Inf)
```
read database

```{r}
setwd("C:/Users/tbao/Desktop/CUNY MSDS notes/607/week 6/project 2/New York City Leading cause of Death")
getwd()
df <- read.csv("C:/Users/tbao/Desktop/CUNY MSDS notes/607/week 6/project 2/New York City Leading cause of Death/New_York_City_Leading_Causes_of_Death.csv", header=TRUE, sep=",")
datatable(df)
```
leading causes of death each year for men and women.
```{r}
df_death <- df %>% subset(select = c(Sex,Year,Cause.of.Death,Count)) %>% 
  group_by(Sex,Year) %>%
  filter(Count == max(Count)) %>% distinct(Sex,Year,Cause.of.Death,Count)
head(df_death)
```


Display the leading causes of death each year for each ethnic group.
```{r}
df_death_eth <- df %>% subset(select = c(Ethnicity,Year,Cause.of.Death,Count)) %>% 
  group_by(Ethnicity,Year) %>%
  filter(Count == max(Count)) %>% distinct(Ethnicity,Year,Cause.of.Death,Count)

head(df_death_eth)
```
Calculate which cause of death has reduced the most and which has increased the most in the years given
```{r}
df_death_trends <- df %>% subset(select = c(Year,Cause.of.Death,Count)) %>%
  group_by(Year, Cause.of.Death, the_min = min(Count), the_max = max(Count), the_diff = (the_max - the_min)) %>% count(Year, Cause.of.Death, max(the_diff))

datatable(df_death_trends)
```
Calculate which cause of death has remained stable over the years given.
```{r}
df_death_stable <- df %>% subset(select = c(Year,Cause.of.Death,Count)) %>%
  group_by(Year, Cause.of.Death, the_min = min(Count), the_max = max(Count), the_diff = (the_max - the_min)) %>% count(Year, Cause.of.Death, min(the_diff))

datatable(df_death_stable)
```

