---
title: "Jun Pan DATA 607 Project 2 3rd database"
author: "Jun Pan"
date: "October 6, 2018"
output: html_document
---

United Nations Development Program sets "Zero Hunger" as one of their sustainable development goals (SDGs). The prevalence of undernourishment (% of population) under code SN.ITK.DEFC.ZS are evaluated in the database.  

Database source: https://datacatalog.worldbank.org/dataset/sustainable-development-goals

Set environment
```{r}
library(dplyr)
library(tidyr)
library(knitr)
library(ggplot2)
library(stringr)
```
```{r}
read.csv("C:/Users/tbao/Desktop/CUNY MSDS notes/607/week 6/project 2/World Bank/SDGData.csv")
```
It shows that the database has 91524 rows and 33 columns

```{r}
sdgdata = read.table("C:/Users/tbao/Desktop/CUNY MSDS notes/607/week 6/project 2/World Bank/SDGData.csv", header=TRUE, sep=",", na.strings = c("", "NA"))
sdgdata$X <- NULL
```
```{r}
head(sdgdata)
```
```{r}
tail(sdgdata)
```

change the names of the columns
```{r}
names(sdgdata)[1] <- "country_name"
names(sdgdata)[2] <- "country_code"
names(sdgdata)[3] <- "indicator_name"
names(sdgdata)[4] <- "indicator_code"
```

```{r}
head(sdgdata)
```

Check how many countries in the database
```{r}
sdgdata %>% dplyr::select(country_name) %>% dplyr::n_distinct()
```
list names of all 263 countires
```{r}
countries <- unlist(sdgdata %>% dplyr::select(country_name) %>% dplyr::distinct())
names(countries) <- NULL
rownames(countries) <- c()
kable(countries, format="markdown")
```


Create a subset for indicator  Prevalence of undernourishment (% of population) under code SN.ITK.DEFC.ZS
For the years before 2000 and after 2015, we do not have the data.  So we are going to exclude those columnes inour subset of database.  

```{r}
sdg_subset <- sdgdata %>% dplyr::filter(sdgdata$indicator_code == "SN.ITK.DEFC.ZS")
head(sdg_subset)
sdg_subset$X1990 <- NULL
sdg_subset$X1991 <- NULL
sdg_subset$X1992 <- NULL
sdg_subset$X1993 <- NULL
sdg_subset$X1994 <- NULL
sdg_subset$X1995 <- NULL
sdg_subset$X1996 <- NULL
sdg_subset$X1997 <- NULL
sdg_subset$X1998 <- NULL
sdg_subset$X1999 <- NULL
sdg_subset$X2016 <- NULL
sdg_subset$X2017 <- NULL

head(sdg_subset)
```
```{r}
dim(sdg_subset)
```
keep only Completed cases in subset
```{r}
sdg_subset <- sdg_subset[complete.cases(sdg_subset), ]
```

```{r}
dim(sdg_subset)
```
Tidy subset:(1)drop indicator_name and indicator_code columes; (2)combine the columes from X2000 to X2015 to a new columes with the name of "year"
```{r}
sdg_subset_transform1 <- tidyr::gather(sdg_subset, "year", "undernourishment_SN.ITK.DEFC.ZS", 5:20)

sdg_subset_transform1$indicator_code <- NULL
sdg_subset_transform1$indicator_name <- NULL

head(sdg_subset_transform1)
```



Calculate the mean for all countries for each year from 2000 to 2015.
```{r}
sdg_meanPerYear <-
  sdg_subset_transform1 %>% 
  dplyr::group_by(year) %>% 
  dplyr::summarise(mean = mean(undernourishment_SN.ITK.DEFC.ZS))

for (i in 1:nrow(sdg_meanPerYear)) {
  sdg_meanPerYear[i,1] <- stringr::str_extract(sdg_meanPerYear[i,1], "[0-9]{4,4}")
}

print(sdg_meanPerYear)
```

Calculate the mean prevalence of undernourishment for each country from year 2000 to 2015.
```{r}
sdg_meanPerCountry <-
  sdg_subset_transform1 %>% 
    dplyr::group_by(country_name) %>%  
    dplyr::summarise(country_mean = mean(undernourishment_SN.ITK.DEFC.ZS))

names(sdg_meanPerCountry)[1] <- "country_name"

for (i in 1:nrow(sdg_meanPerYear)) {
  sdg_meanPerYear[i,1] <- stringr::str_extract(sdg_meanPerYear[i,1], "[0-9]{4,4}")
}

print(sdg_meanPerCountry)
```
Calculate the mean prevalence of undernourishment of 214 countries (yr 2000 to yr 2015)

```{r}
sdg_meanPerCountry
```
Summary statistics for the mean prevalence of all countries
```{r}
unlist(sdg_meanPerCountry %>% dplyr::summarise(mean=mean(country_mean), 
                                median=median(country_mean),
                                IQR = IQR(country_mean),
                                sd = sd(country_mean),
                                min = min(country_mean),
                                max = max(country_mean)))
```

mean prevelance of undernourishment per country

```{r}
barfill <- "#FF6666"
barlines <- "#1F3552"

e <- ggplot(sdg_meanPerCountry, aes(x = country_mean)) +
        geom_histogram(aes(y = ..count..), binwidth = 1,
                       colour = barlines, fill = barfill) +
        scale_x_continuous(name = "Mean prevalence of undernourishment from year 2000 to 2015",
                           breaks = seq(0, 50, 5),
                           limits=c(0, 60)) +
        scale_y_continuous(name = "Count") +
        ggtitle("Mean Prevalence of Undernourishment Per Country")
e
```
mean prevalence of undernourishment from year 2000 to 2015
```{r}
sdg_meanPerYear
```

```{r}
f <- ggplot() + geom_bar(aes(y = mean, x = year, fill = "#1F3552"), data = sdg_meanPerYear, stat = "identity") + guides(fill=FALSE)
f
```
Rank countries of prevelance of undernourishment in descending order
```{r}
sdg_meanPerCountry %>% dplyr::arrange(desc(country_mean))
```

